---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


# urbioconnect

<!-- badges: start -->
[![R-CMD-check](https://github.com/urbio-ecology/urbioconnect/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/urbio-ecology/urbioconnect/actions/workflows/R-CMD-check.yaml)
[![R-universe version](https://urbio-ecology.r-universe.dev/urbioconnect/badges/version)](https://urbio-ecology.r-universe.dev/urbioconnect)
[![Codecov test coverage](https://codecov.io/gh/urbio-ecology/urbioconnect/graph/badge.svg)](https://app.codecov.io/gh/urbio-ecology/urbioconnect)
<!-- badges: end -->

`urbioconnect` implements methods that help quantify the ecological connectivity for different urban wildlife species.

This is done by providing wildlife **habitat** and urban **barrier** information, the roaming range (in metres) of the wildlife, and some other parameters.

For example, raster grid data on blue tongued-lizard **habitat**, how far their roaming range is (say, 100 metres) and raster grid data on road and buildings (**barriers**). It then computes metrics that help assess how connected these habitat patches are:

* effective mesh size
* probability of connectedness
* number of patches, 
* total patch area and 
* mean patch area

Note that you can use **vector** or **raster** data for habitat and barrier formats.

This method is described in:

> Kirk, H., Soanes, K., Amati, M., Bekessy, S., Harrison, L., Parris, K., Ramalho, C., van de Ree, R., & Threlfall, C. (2023). Ecological connectivity as a planning tool for the conservation of wildlife in cities. *MethodsX*, 10, 101989. <https://doi.org/10.1016/j.mex.2022.101989>

See the [getting started](https://urbio-ecology.github.io/urbioconnect/articles/getting-started.html) vignette for more details.

## Shiny app

We include a Shiny app for interactive analysis and report generation. You can launch the interactive Shiny app with:

```r
run_connectivity_app()
```

A hosted version is available at: <https://njtierney.shinyapps.io/urbioconnect/>

However, it is worth noting you do not need the shiny app to perform analyses, all functions have been designed to work together in a pipeline - see example usage below.

## Installation

Install from R-universe:

```r
install.packages(
  "urbioconnect",
  repos = c("https://urbio-ecology.r-universe.dev", "https://cloud.r-project.org")
)
```

Or install from GitHub:

```r
# install.packages("pak")
pak::pak("urbio-ecology/urbioconnect")
```

## Get started

```{r get-started}
library(urbioconnect)

# load example habitat and barrier rasters
habitat <- example_habitat()
barrier <- example_barrier()

# run the full raster pipeline at a 10m buffer distance
areas <- habitat_connectivity(
  habitat  = habitat,
  barrier  = barrier,
  distance = 10
)

areas
```

Summarise the connectivity metrics:

```{r summarise}
summarise_connectivity(
  area_squared     = areas$area_squared,
  area_total       = areas$area,
  buffer_distance  = 100,
  target_resolution = 500,
  data_resolution  = 10,
  aggregation_factor = 50,
  species_name     = "Blue-tongued Lizard"
)
```


## Use raster and vector

You can use raster and vector data with `urbioconnect`:

| | Raster pipeline | Vector pipeline |
|---|---|---|
| Input format | `terra` SpatRaster | `sf` polygons |
| Main function | `habitat_connectivity()` | `sf_habitat_connectivity()` |
| Best for | Large study areas, GeoTIFF output | Small precise study areas, exact polygon boundaries |

You can see the vignette, ["raster-vs-vector"](https://urbio-ecology.github.io/urbioconnect/articles/raster-vs-vector.html) for a side-by-side comparison and guidance on which to choose.

# Example usage

We have an example pipeline using `targets` and `geotarget`, which includes generating reports as output at https://github.com/urbio-ecology/urbio-eco-targets. See the vignette ["Using urbioconnect in a targets pipeline"](https://urbio-ecology.github.io/urbioconnect/articles/targets-pipeline.html) for more detail.

# Acknowledgements

We would like to thank Kylie Soanes, Marco Amati, Sarah Bekessy, Lee Harrison, Kirsten Parris, Cristina Ramalho, Rodney van de Ree, and Caragh Threlfall for their work on the original paper this work is based upon. We would also like to thank Hugh Stanford, Nadine Gaskell, Kerryn Kneebone, and Nicholas Golding for their comments and insight during the development of the methods and software.
