---
title: "Brief Exploration of results"
author: "Nicholas Tierney"
date: "`r format(Sys.time(), '%d %B, %Y')`"
format: 
  html:
    code-fold: true
    embed-resources: true
    lightbox: true
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
## target knits qmds in their own session, so load libraries here.
source(here::here("packages.R"))
tar_source() # not always necessary
```

```{r}
#| label: load-targets
#| include: false
with_dir(here(), {
  tar_load(habitat_raster)
  tar_load(barrier_raster)
  tar_load(buffered_habitat)
  tar_load(patch_id_raster)
  tar_load(areas_connected)
  tar_load(results_connect_habitat)
  tar_load(buffer_distance)
  tar_load(species_name)
  tar_load(saved_spatraster)
})
```

```{r}
urbio_pal <- scico::scico(n = 11, palette = "tofino")
urbio_pal_cut <- urbio_pal[c(6:11)]
# urbio_pal_cut |> swatchplot()
# habitat  = 3/4
# buffer = 6
# barrier = 1/2

urbio_cols <- list(
    habitat = urbio_pal_cut[2],
    buffer = urbio_pal_cut[5],
    barrier = col2hex("white")
)

## TODO for mdsumner - can we get a buffer/outline of the city boundaries
```

## Habitat, Buffered Habitat, and Barrier

```{r}
#| label: plot-barrier-habitat-buffer
#| panel: tabset
#| output: asis

buffer_barrier_plots <- map2(
  .x = buffered_habitat,
  .y = buffer_distance,
  .f = function(buffered_habitat, buffer_distance) {
plot_barrier_habitat_buffer(
  barrier = barrier_raster,
  buffered = buffered_habitat,
  habitat = habitat_raster,
  distance = buffer_distance,
  species_name = species_name,
  col_barrier = urbio_cols$barrier,
  col_buffer = urbio_cols$buffer,
  col_habitat = urbio_cols$habitat,
  col_paper = "grey96"
) 
  }
) |> 
  setNames(buffer_distance)

show_tabs(buffer_barrier_plots, "Buffer: ")
```

## Patch ID

```{r}
#| label: plot-patches
#| panel: tabset
#| output: asis

patch_plots <- map2(
  .x = patch_id_raster,
  .y = buffer_distance,
  .f = \(x, y) plot_patches(patch_id = x, distance = y)
  ) |> 
  setNames(buffer_distance)

show_tabs(patch_plots, message = "Buffer: ")
```

# Area and patch information for each buffer

```{r}
#| label: display-areas

summary_dts <- areas_connected |> 
  setNames(buffer_distance) |> 
  bind_rows(
    .id = "buffer"
  ) |> 
  datatable()

summary_dts
```

## Prob connectedness and summary information for each buffer

```{r}
results_connect_habitat |> datatable()
```

## Longer: Prob connectedness and summary information for each buffer

```{r}
results_connect_habitat |> 
pivot_longer(
    cols = -c(species_name,
              buffer_distance)
  ) |> 
  datatable(
    options = list(
      pageLength = 8L
    )
  )
```

## Visualisation of changes in key stats over buffer distance

```{r}
#| fig-height: 10
plot_connectivity(results_connect_habitat)
```


## Running time

It is useful to understand how long the pipeline takes, let's take a look at the timing of some key targets

```{r}
#| label: running-time
target_timings <- tar_meta() |> dplyr::select(name, bytes, seconds)

target_timings |>
  drop_na() |>
  arrange(desc(seconds)) |>
  mutate(
    bytes_pretty = pretty_bytes(bytes),
    .after = bytes
  ) |>
  rename(
    target_name = name,
    runtime_seconds = seconds
  ) |>
  datatable()
```
