---
title: "Demonstrate Urban Connectivity example with Fairy Wren"
format: 
  html:
      toc: true
---

This is an analysis of how connected the habitat of the Superb Fairy Wren is in a region of Victoria, Knox. We use a vector based (as opposed to raster) geospatial data, in this first instance. Our goal is to provide metrics of connectedness, as well as maps showing the connections in the landscape.

First we have to source the R packages and the R files in the R directory:

```{r}
#| echo: false
# prep habitat layer (lidar and shrubs)
## vector based
source(here::here("packages.R"))
dir_map(path = here::here("R/"), fun = source)
```

This will eventually be an R package, so just noting that here.

We load two shapefiles:

1. The barrier: Roads, houses, etc
2. The habitat: Grass, trees, water, etc

Out inputs are essentially the movements of species (habitat), against the barriers in the city (barriers). Our outputs are a map of connectedness, and various metrics of connectedness.

Reading the barrier for the superb fairy wren, this is all main roads:

```{r}
barrier <- read_geometry(here("data/allSFWRoads.shp"))

plot(barrier, border = "orange2")
```

We can then read in the habitat of the superb fairy wren. Here it is all understorey from the LiDATa data. We also run `clean` to clean up the edges of the shapefile data, to rmove some of the very high level details that we do not need:

```{r}
habitat <- read_geometry(here("data/superbHab.shp")) |> clean()
plot(barrier, border = "orange2")
plot(habitat, border = "midnightblue", add = TRUE)
```

We can then put these inputs into the `habitat_connectivity()` function, which returns to us a data frame that has identified the various patches that are created by the barriers, as well as the areas of each of these patches

```{r}
connect_habitat <- habitat_connectivity(
  habitat = habitat,
  barrier = barrier,
  distance = 250
)

connect_habitat
```

We can can summarise these metrics with `summarise_connectivity()`:

```{r}
results <- summarise_connectivity(connect_habitat)
results
```

There are a few steps that go on behind the scenes in `habitat_connectivity()` - below we step through each of these steps and, where relevant, show the output of the map.

First we create a 250 buffer around the environment, which tells us that the fairy wren have a range of 500m. We can see the habitat in darker green, and the buffer in lighter green.

```{r}
# buffer the habitat by distance
buffer <- habitat_buffer(habitat, distance = 250)
```


```{r}
plot(habitat, border = "forestgreen")
plot(
  buffer,
  border = "darkgreen",
  add = TRUE,
  col = alpha(colour = "forestgreen", alpha = 0.25)
)
```

We then essentially fragment the environment by the barriers with `fragment_geometrc()`:

```{r}
# create fragmentation geometry
fragment <- fragment_habitat(buffer, barrier = barrier)

plot(habitat, border = "darkgreen", col = "forestgreen")
plot(
  fragment$fg,
  border = "transparent",
  col = alpha(colour = "midnightblue", alpha = 0.5),
  add = TRUE
)
```

Then we remove all habitat that lies under the barriers, and identify the patches that are connected and not severed by the barriers.

```{r}
# remove all habitat under barriers
remaining_habitat <- drop_habitat_under_barrier(habitat, barrier = barrier)

# identify remaining habitat patches according to which connected area they
# belong to
id_remaining_habitat <- assign_patches_to_fragments(remaining_habitat, fragment)
# id_remaining_habitat is a key output that shows you the connected areas in a
# landscape
id_remaining_habitat
```

The dataset contains the polygons of each patch - we then get add the area of each patch, and group by each patch to get the total area of each patch as well as the area squared:

```{r}
# calculate area of each habitat patch
area_hectares <- add_patch_area(id_remaining_habitat)
area_hectares
# group the patches by connected area ID
connect_habitat <- aggregate_connected_patches(area_hectares)
connect_habitat
```

This is the same as our `habitat_connectivity()` function:

```{r}
# or, as one step
connect_habitat2 <- habitat_connectivity(
  habitat = habitat,
  barrier = barrier,
  distance = 250
)

all.equal(connect_habitat, connect_habitat2)
```

We then summarise these measurements of connectivity:

```{r}
# calculation
results <- summarise_connectivity(connect_habitat)
results
```

